Social Visual Analytics: DevOps for Data Scientists

### 2015-03-25 introduction.tex
summarize interviews

Interviews in {\it Data Scientists at Work} by Sebastian Gutierrez illuminate
the teamwork and sharing in the current practice of visual analytics in industry.
We particularly appreciated the remarks by Jonathan Lenaghan about the potential
for continuous integration and publication.

% \begin{addmargin}[1em]{1em}  % left and right margin change
% part1.txt
Erin Shellman [of Nordstrom] described the benefits of
continuous deployment of experiments, and sharing knowledge
through software and data analysis artifacts.
``Finally, prototyping our products so that internal
customers can use them early on has been crucial for
our success. It doesn't even have to be something
fancy. For instance, our recommendations preview
tool doesn't have a particularly interesting
visualization, but its enough to show the result of our
algorithms. Now we can shoot off a URL to internal
customers and it allows them to sit at their desk and
understand the behavior of our product and
experiment with it, and provide feedback way before
we're talking about getting it into production. This
has been super helpful and has been a really great
way to get people excited about what we're working
on...
%Building and maintaining those relationships are
%just like anything else-- you always have to be
%working on them. Nurturing those relationships are
%part of the job, and if you leave them unattended,
%they might not be there later.
%
% part2.txt
%Gutierrez: How do you share the knowledge you are
%building with others?
We’re obsessed with Confluence from
Atlassian, and the Data Lab has a very active
Confluence space. Recommendo is fully documented
on Confluence, so anyone in the company could learn
how it’s built, where to find it, and how to use it.
We also share exploratory analyses and reports on
Confluence so that we can still exchange knowledge
even if the work didn't make it into a larger project.''

% part3.txt
John Foreman commented on the scope of the work in
data science teams.
``I lead the data science team at
MailChimp, and I like to get my hands dirty too.
Some of the big pieces of my day involve working
with my team to take stock of current projects and
figure out where to go next, doing my own work and
prototyping things-- I do projects just like my peers
do and then also my talking with other teams,
talking with management, and planning for the
future.
On our team, we've got different folks facing different
% part4.txt
kinds of projects. We've got one person who really
owns compliance and looks at our compliance
processes. We've got another data scientist who
focuses on more of the user experience side of the
house and understanding our customers. I help them
and others as is needed while I do the three things
mentioned earlier-- build data products, be a
translator, and have conversations with the data
science community.''

% part5.txt
Jonathan Lenaghan from PlaceIQ discussed to the potential
for continuous deployment in data science.
``If you look at dev ops right now, they have things
such as continuous integration, continuous build,
automated testing, and test harnesses-- a1l of which
map very well from the dev ops world to the data ops
(a phrase I stole from Red Monk) world very easily. I
think this is a very powerful notion. It is important to
have testing frameworks for all of your data, so that if
you make a code change, you can go back and test all
of your data.''

% part6.txt
Anna Smith from Bitly spoke about the benefits of access to
a rich environment of shared data and code.
``In addition to Bitly's extraordinary data sets,
I was given all the resources that I needed. These were
resources that I’d only heard about but had never
seen. I went from asking what a Hadoop cluster was,
as my school didn't have one, to being able to work
with one. Now I could play with one to figure out how
to make my work better and how to change my
algorithms so they were cleaner and ran faster on
Hadoop. The whole transition was eye opening. Not
only that, I could also look at other people’s code and
play with their code and data sets as well. It was
fantastic and I was convinced that working with data was my thing.''
% \end{addmargin}

% (Hiring ``full stack developers'' to perform integration seems to be a trend, but does not achieve the potential benefits of specialization and division of labor in larger teams, and is not a practice to be encouraged.)

Unfortunately, processes and technology that support visual analytics today
work are fragmented.
Knowledge is shared through informal conversations, email online messages,
meetings, phone calls, within code itself, in wikis, in project documents and
by other means. Results of experiments are often shared first by copying static
output, later, by holding meetings that lead to moving or even rewriting
entire programs.

This disjointed, inconsistent process inhibits the use of tools to bridge
the many gaps and to help people find resources.
For example, even when code is shared, we still not may know where to
find examples that show how to use it.
When we find a useful data source, we don't know what packages or
other data sources usually go along with it, what problems with the data
others have already discovered and corrected, or where to obtain online feeds
to get updates.

This situation definitely affects how well data science teams work.
It limits our ability to share work and to rapidly deploy and refine results.
We have seen this in our lab environment.
Other authors report similar findings.

Some symptoms of the current situation are:
\begin{itemize}
\item It is hard to find data, metadata, and knowledge about them.
\item Most coordination is done through meetings, whose content is not related to other artifacts and may not even be stored.
\item Production deployment requires porting code to a different environment or even completely rewriting it, thwarting continuous release.
\item Many tools adopt a standalone or single developer perspective,
not suited to collaboration and web deployment. 
\item Analysts find insufficient support for tracing events or issues from production back to the EDA process.
\end{itemize}

>>>>

%% Interviews in {\it Data Scientists at Work} by Sebastian Gutierrez illuminate
%% the teamwork and sharing in the current practice of visual analytics in industry.
%% We particularly appreciated the remarks by Jonathan Lenaghan about the potential
%% for continuous integration and publication.

%% % \begin{addmargin}[1em]{1em}  % left and right margin change
%% % part1.txt
%% Erin Shellman [of Nordstrom] described the benefits of
%% continuous deployment of experiments, and sharing knowledge
%% through software and data analysis artifacts.
%% ``Finally, prototyping our products so that internal
%% customers can use them early on has been crucial for
%% our success. It doesn't even have to be something
%% fancy. For instance, our recommendations preview
%% tool doesn't have a particularly interesting
%% visualization, but its enough to show the result of our
%% algorithms. Now we can shoot off a URL to internal
%% customers and it allows them to sit at their desk and
%% understand the behavior of our product and
%% experiment with it, and provide feedback way before
%% we're talking about getting it into production. This
%% has been super helpful and has been a really great
%% way to get people excited about what we're working
%% on...
%% %Building and maintaining those relationships are
%% %just like anything else-- you always have to be
%% %working on them. Nurturing those relationships are
%% %part of the job, and if you leave them unattended,
%% %they might not be there later.
%% %
%% % part2.txt
%% %Gutierrez: How do you share the knowledge you are
%% %building with others?
%% We’re obsessed with Confluence from
%% Atlassian, and the Data Lab has a very active
%% Confluence space. Recommendo is fully documented
%% on Confluence, so anyone in the company could learn
%% how it’s built, where to find it, and how to use it.
%% We also share exploratory analyses and reports on
%% Confluence so that we can still exchange knowledge
%% even if the work didn't make it into a larger project.''

%% % part3.txt
%% John Foreman commented on the scope of the work in
%% data science teams.
%% ``I lead the data science team at
%% MailChimp, and I like to get my hands dirty too.
%% Some of the big pieces of my day involve working
%% with my team to take stock of current projects and
%% figure out where to go next, doing my own work and
%% prototyping things-- I do projects just like my peers
%% do and then also my talking with other teams,
%% talking with management, and planning for the
%% future.
%% On our team, we've got different folks facing different
%% % part4.txt
%% kinds of projects. We've got one person who really
%% owns compliance and looks at our compliance
%% processes. We've got another data scientist who
%% focuses on more of the user experience side of the
%% house and understanding our customers. I help them
%% and others as is needed while I do the three things
%% mentioned earlier-- build data products, be a
%% translator, and have conversations with the data
%% science community.''

%% % part5.txt
%% Jonathan Lenaghan from PlaceIQ discussed to the potential
%% for continuous deployment in data science.
%% ``If you look at dev ops right now, they have things
%% such as continuous integration, continuous build,
%% automated testing, and test harnesses-- a1l of which
%% map very well from the dev ops world to the data ops
%% (a phrase I stole from Red Monk) world very easily. I
%% think this is a very powerful notion. It is important to
%% have testing frameworks for all of your data, so that if
%% you make a code change, you can go back and test all
%% of your data.''

%% % part6.txt
%% Anna Smith from Bitly spoke about the benefits of access to
%% a rich environment of shared data and code.
%% ``In addition to Bitly's extraordinary data sets,
%% I was given all the resources that I needed. These were
%% resources that I’d only heard about but had never
%% seen. I went from asking what a Hadoop cluster was,
%% as my school didn't have one, to being able to work
%% with one. Now I could play with one to figure out how
%% to make my work better and how to change my
%% algorithms so they were cleaner and ran faster on
%% Hadoop. The whole transition was eye opening. Not
%% only that, I could also look at other people’s code and
%% play with their code and data sets as well. It was
%% fantastic and I was convinced that working with data was my thing.''
%% % \end{addmargin}

%% % (Hiring ``full stack developers'' to perform integration seems to be a trend, but does not achieve the potential benefits of specialization and division of labor in larger teams, and is not a practice to be encouraged.)

Unfortunately, the processes and technologies supporting data analysis visual analytics are fragmented.
Knowledge is shared through informal conversations, emails,
meetings, phone calls, source code itself, in wikis, in project documents and
by other means. Results of experiments are often shared first by copying static
output (usually a simple screen shot or image output). By the time
decisions are made based on the results of that screenshot, the data
and the processes might have changed enough to bring to question the
original decision.

%% This disjointed, inconsistent process inhibits the use of tools to bridge
%% the many gaps and to help people find resources.
%% For example, even when code is shared, we still not may know where to
%% find examples that show how to use it.
%% When we find a useful data source, we don't know what packages or
%% other data sources usually go along with it, what problems with the data
%% others have already discovered and corrected, or where to obtain online feeds
%% to get updates.

This situation can affect how data science teams; as analysts and
toolbuilders, we have experienced this first-hand in our environment, and
others report similar findings. Gutierrez collected interviews with data
scientists~\cite{Gutierrez}, which include the following:

\begin{itemize}
\item (upon being given access to other code and data analysis, in
order to learn about Hadoop) ``[...] Not only that, I could also look at other people’s code and
play with their code and data sets as well''
\item (discussing the value of sharing prototypes rather than static
data) ``[...] Finally, prototyping our products so that internal
customers can use them early on has been crucial for
our success. [...] Now we can shoot off a URL to internal
customers and it allows them to [...] provide feedback way before
we're talking about getting it into production.''
\item (on sharing more than just a finished product) ``We also share
exploratory analyses and reports [...] so that we can still exchange
knowledge even if the work didn't make it into a larger project.''
\item (on changing analyses and processes) ``It is important to
have testing frameworks for all of your data, so [...] you can go back and test all
of your data. 
\end{itemize}

Currently, these concerns need to be addressed \emph{alongside} the
visual analytics environment. Some symptoms of the current situation are:
\begin{itemize}
\item It is hard to find data, metadata, and knowledge about them.
\item Most coordination is done through meetings, whose content is not related to other artifacts and may not even be stored.
\item Production deployment requires porting code to a different environment or even completely rewriting it, thwarting continuous release.
\item Many tools adopt a standalone or single developer perspective,
not suited to collaboration and web deployment. 
\item Analysts find insufficient support for tracing events or issues from production back to the EDA process.
\end{itemize}
